Postings during the next few days quickly reduced our numbers in depot
company and each morning I expected to hear my name read out from the
daily posting list but several weeks were to elapse before this
happened. Meanwhile, with an ever dwindling number of sappers, I
paraded each morning a 08.00 hours to answer roll call, listen to the
posting list and be detailed to fatigues. \CsergeantM
Richards who took this parade had a wonderfully deceptive way of
getting men to volunteer for unwanted fatigues. Alec Middleton who
slept in our room was caught in this deceptive way to volunteer for
`fire picket'. The way \sergeantM Richards phrased his needs was to
make one believe that you were volunteering to become a member of the
barracks' fire brigade, with the added probability of a permanent
posting at Fulwood. Alec was friendly with a girl in Preston and a
permanent posting here would have suited him very well. While the men
on parade were looking at each other and wondering where the catch was
and Alec was dithering about volunteering for the job, I tried to
convince him to chance it saying that perhaps this was the opening he
wanted. A sapper on his other side joined in to egg him on. Finally
Alec, mumbling that it was against his nature to volunteer for
anything in the army, stepped forward and \sergeantM Richards told him to
report to company office at 18.00 hours for further details. Alec as
usual dressed up to spend the evening with his girlfriend in Preston
and called at Company Office on his way out of barracks.

He was soon back in a furious temper. He wasn't going to be a member
of the barrack's fire brigade but had unwittingly volunteered for fire
picket duty and was now confined to barracks for a week. The names he
called me and Johnson, the other sapper who had egged him on, were not
complimentary. Alec who was a single man and appeared to have a
comfortable flow of cash, paid someone each evening to stand in for
him on the fire picket so that he could get into Preston.

I had become friendly with a regular soldier in our room who was on
the barracks permanent staff. He was an army-trained tailor and spent
his time in the tailor's shop altering and repairing uniforms for the
officers of the establishment and for a few shillings was ready to do
stitching for anyone else. He altered my battle dress blouses for
me. He was almost ready for discharge when war broke out and so he had
a wide experience of army life. From him I learned many tips on how
to make life in the army more comfortable. This knowledge could not be
found in any training manual but put together it would have made a
nice little booklet called, `A Scrounger's Guide to the Army'.

On most parades I was detailed to draw a carpenter's toolkit and sent
to repair windows and doors around the barracks, work which suited me.
My best fatigue was when I was sent to the records office in Preston
to alter shelving and storage racks under guidance of the
A.T.S. officer in charge there. They appeared to have an unlimited
supply of tea and biscuits. Having to have my dinner at the barracks
meant that the time of four bus journeys a day left little time for
work at the office and using those little bits of knowledge gained
from my tailoring friend, I made this job last a long time, so much so
that \sergeantM Richards jokingly suggested that I should find billets
there. The price of my bus fares were refunded by company office on
the production of my tickets.

Scrounging about the barracks and making jobs last as long as possible
became very boring. The familiar faces of the men I had trained with
had all disappeared and I was beginning to feel out on a limb, so I
decided to have a word with the postings clerk. Seeing him alone in
the postings office one evening when I had walked that way to read
daily orders, I went inside and asked him if he knew of any reason why
I hadn't been posted. He asked my name, looked through his files and
told me that as far as he knew there was nothing holding me at
Fulwood. That was a relief anyway for I wasn't happy with the place.
When he told me that sappers from here were sometimes sent to Kineton
I just prayed that I could get onto that kind of posting and I had
visions of cycling home from Kineton to Kenilworth at weekends. My
dreams came to an abrupt end soon afterwards for I was posted to
Darlington where a new artisan works company was being formed.

